%!TEX root = main.tex

\documentclass[../main.tex]{subfiles}

\begin{document}

\chapter*{Acknowledgements}
I would like to thank Prof. Gary Scavone for serving as my thesis advisor. His knowledge of physical modeling, acoustics and acoustic measurement techniques served as a invaluable sounding board during the development process. Many thanks to Prof. Ichiro Fuginaga for agreeing to be my thesis examiner. Laurent Olivier-Lord, Andrew David and John Gilbert deserve credit in helping me edit my French abstract. Dr. Andre Van Schyndel provided valuable feedback to help fine tune the thesis.

Many thanks to both Aybar Adin and Vlad Baran of McGill's Sound Recording Program as well as Yves Méthot of CIRMMT. All of these people were extremely helpful in determining which microphones and other equipment to use for the physical measurements. Song Wang was also extremely helpful in brainstorming ideas when attempting to devise a measurement system to capture the decay parameters associated with longitudinal string motion.

On the DSP side of things, I would like to thank Max Henry as well as inhabitants of the \emph{\#dsp} channel on The Audio Programmer Discord Server. Their help in exploring DSP concepts, listening to my results, as well as debugging the nitty-gritty implementation details related to things like time-varying Lagrange interpolation benefited me greatly.

Finally, I would like to thank Nadia Stevens for all her help, support and love throughout this process. This includes picking up the slack around the apartment when I had deadlines to meet as well as when I was operating at a reduced capacity due to a hard to diagnose medical ailment which decided to appear when a healthcare crisis was happening across the country.
\end{document}